A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-08-13 paper/fascd.tex
finish the sweep of final polish edits; an EB remains ...

\usepackage{todonotes}
\newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
%\newcommand{\elb}[1]{\todo[inline, color=blue!20]{EB: #1}}

@@

@@

A key choice in devising a multilevel scheme for VIs is how to construct coarser-mesh bounds.
Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} derived from an admissible finest-level iterate.  The next Example illustrates why we do \emph{not} apply injection operators directly\footnote{An exception occurs in FMG; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.

\begin{example}  \label{ex:directRbad}  
Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).  Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ strictly less than the maximum of $\underline{\gamma}^1$.  Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.  For admissible boundary conditions the fine-level constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.  However, various schemes to coarsen the constraints by direct transfers of $\underline{\gamma}^1,\overline{\gamma}^1$ are doomed to failure.  Suppose we propose
\begin{equation}
    \label{eq:fe:badstandard}
    \underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}
\end{equation}
Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.  In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($w^0\in\mathcal{K}^0$; $\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level ($Pw^0 \notin \mathcal{K}^1$), and attempting to fix this by projecting into $\mathcal{K}^1$ would introduce high frequencies not present in $w^0$.  On the other hand, if
\begin{equation}
    \label{eq:fe:badmonotone}
    \underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}
\end{equation}
then $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.  However, now the coarse-level constraint set $\mathcal{K}^0$ is empty because $\underline{\gamma}^0 > \overline{\gamma}^0$ (Figure \ref{fig:directRbad}).
\end{example}

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\input{tikz/directRbadfine.tex}
&
\input{tikz/directRbadcoarse.tex}
\end{tabular}
\caption{For the finer-level constraints $\underline{\gamma}^1$ (blue) and $\overline{\gamma}^1$ (red) on the left, direct application of $\iR,\minR,\maxR$ is problematic.  For example, if we apply the monotone injections then the resulting coarser-level constraint set $\mathcal{K}^0$ is empty (right).  Defect constraints avoid this difficulty.}
\label{fig:directRbad}
\end{figure}

@@

This construction of multilevel box constraints, likewise for the unilateral construction in \cite{GraeserKornhuber2009}, permits large, yet admissible, coarse corrections, which promotes multilevel solver efficiency.  However, when coarsening the problem downward in a V-cycle, one cannot predict the upcoming coarse corrections and so one must construct constraint sets from which any sum will be admissible.  With this in mind, and looking forward to telescoping sums \eqref{eq:fe:telescoping}, we compute the LDC differences \cite{GraeserKornhuber2009}:
\begin{equation}
\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \qquad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}.  \label{eq:fe:philevels}
\end{equation}
(Take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  By definition, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and etc.)  Note that while $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, they are not ordered as in \eqref{eq:fe:chiordering}.

@@

Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute corrections in $\mathcal{D}^j$ and $\mathcal{U}^j$.  For general box constraints the upward direction is, however, based on \emph{incomplete} CDs as we now explain.  Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, though this becomes a true CD satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} in unilateral cases.  The following Lemma, apparently new, extends the unilateral CD construction underlying Algorithm 4.2 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.

@@

In unilateral cases we can also construct decomposition operators $\Pi_i$ and thereby show \eqref{eq:constraintrestrictionsum}.  For concreteness suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels; the $\underline{\gamma}^J=-\infty$ case is similar.  For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) injection $j-i$ times, mapping into $\mathcal{V}^i$: $I_{j\to i}^\ominus v = \minR(\cdots(\minR v)\cdots) $.  (Set $I_{j\to j}^\ominus v = v$ and $I_{j\to -1}^\ominus=0$ by definition.)  Now define nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
\begin{equation}
\Pi_i z^j \coloneqq I_{j\to i}^\ominus(z^j - \underline{\chi}^j) - I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j) + \underline{\phi}^i.  \label{eq:fe:unilateraldecompositionoperator}
\end{equation}
(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property \eqref{eq:fe:monotoneproperty} implies $I_{j\to i}^\ominus(z^j - \underline{\chi}^j) \ge I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j)$, thus $\Pi_i z^j \ge \underline{\phi}^i$, so $\Pi_i z^j \in \mathcal{D}^i$, and furthermore stability \eqref{eq:decompositionstability} holds \cite[Theorem 4.2]{GraeserKornhuber2009}.  On the other hand, the following sum telescopes, leaving only its first term plus the sum in \eqref{eq:fe:telescoping}:
\begin{align*}
\sum_{i=0}^j \Pi_i z^j &= z^j - \underline{\chi}^j + \sum_{i=0}^j \underline{\phi}^i = z^j.
\end{align*}
This shows \eqref{eq:constraintrestrictionsum}, thus \eqref{eq:constraintdecomp}, and that \eqref{eq:fe:downwardsuminclusion} is equality.
\end{proof}

One might attempt to convert \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by constructing different decomposition maps $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.  The above proof shows that the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ indeed holds.  However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.  In fact, $\Pi_i$ does not generally map all non-positive elements of $\mathcal{U}^J$ into $\mathcal{D}^i$ if a finite upper LDC is present.  On the other hand, we can not exclude entirely-different strategies for showing full CDs in the general box-constrained case.

\begin{example}  \label{ex:notfullcd}
Let $a > 0$ be an arbitrary positive number.  Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.  Note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.  Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse vertices and value $0$ on the other vertices.  Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^1 \in \mathcal{U}^1$.  However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.  From \eqref{eq:fe:unilateraldecompositionoperator}, $\Pi_1 z^1$ simplifies to $\Pi_1 z^1 = z^1 - \underline{\chi}^1$.  But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.  This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, and $\Pi_1 z^1 \notin \mathcal{D}^1$ would occur.
\end{example}

@@

\begin{lemma}  \label{lem:upwardadmissibility}  For any $j=0,1,\dots,J$ and $0\le k\le j$,
\begin{equation}
\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^{k+1} + \mathcal{U}^k \label{eq:fe:upwardsuminclusion}
\end{equation}
In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, actually $\mathcal{U}^j = \mathcal{U}^k + \sum_{i=k+1}^j \mathcal{D}^i$ as a full CD. \end{lemma}

\begin{proof}  Definition \eqref{eq:fe:philevels} shows that $\underline{\chi}^j = \underline{\phi}^j + \dots + \underline{\phi}^{k+1} + \underline{\chi}^k$ and $\overline{\chi}^j = \overline{\phi}^j + \dots + \overline{\phi}^{k+1} + \overline{\chi}^k$.  If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.  In unilateral cases we may follow the proof of Lemma \ref{lem:downwardadmissibility} with \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but modified in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.  Now the proof goes through as before.
\end{proof}

@@

The V-cycle in this Section extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.2]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}; it reduces to FAS for PDEs when inequality constraints are removed.  Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$, via information from each coarser FE subspace $\mathcal{V}^j$, by approximately solving VI problems in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively; by \eqref{eq:fe:nestedspaces} these corrections are all elements of the fine-level subspace $\mathcal{V}^J$.  The corrections are inductively constructed so that iterates remain admissible at every level.  A key idea is that, by inclusion \eqref{eq:fe:upwardsuminclusion}, up-smoothing solves less-restricted VI problems than down-smoothing.  Going downward one does not know what coarse corrections are coming, thus one must choose smaller constraint sets ($\mathcal{D}^j$) such that the correction sum is admissible once we return to a given level.  Going upward, however, no further coarse corrections can violate admissibility.

Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdcycles}, left).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Relaxation then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for relaxation on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.  The new finest-level iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\input{tikz/vcycle.tex}
&
\input{tikz/fcycle.tex}
\end{tabular}
\caption{Left: The FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets. Right: FMG Algorithm \ref{alg:fascd-fmg} generates each next-finer initial iterate by injection and truncation (doubled edges), then it iterates V-cycles.}
\label{fig:fascdcycles}
\end{figure}

To state the V-cycle algorithm we must specify the problems solved at each mesh level.  As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.  On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:vi}.\footnote{Presumably it uses quadrature etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.}  At the start of each V-cycle we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); they do not depend on the coarse-level corrections.

As we descend we define the initial $j$th-level iterates
\begin{equation}
w^j = \iR(w^{j+1} + y^{j+1}).  \label{eq:fe:definew}
\end{equation}
From $w^j$ we define $\ell^j \in (\cV^j)'$:
\begin{equation}
\ell^j = \begin{cases} \ell^J, & j=J \\
                       f^j(w^j) + R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
\end{equation}
As is typical of FAS \cite[section 5.3]{Trottenbergetal2001}, injection $\iR$ coarsens the (primal) iterates, but canonical restriction $R$ coarsens (dual) linear functionals.  Descending through the levels $j=J,J-1,\dots,1$ we solve the following VI problems for $y^j \in \mathcal{D}^j$:
\begin{equation}
\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j. \label{eq:fe:downvi}
\end{equation}
The initial iterate is $y_{(0)}^j=0$; this is admissible because $\underline{\phi}^j \le 0 \le \overline{\phi}^j$.

@@

However, because the down-smoother has made progress, $y$ can be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the relaxed residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.  These steps yield a problem for the $j$th-level correction $y=y^j \in \mathcal{D}^j$:
\begin{align}
&\ip{f^j\left(\iR(w^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(w^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right)}{v-y^j}, \notag
\end{align}
for all $v\in \mathcal{D}^{j}$.  Though in cluttered form, this is VI problem \eqref{eq:fe:downvi}, once definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} are applied.

>>>>

\usepackage{todonotes}
\newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
\newcommand{\elb}[1]{\todo[inline, color=blue!20]{EB: #1}}

@@

\elb{address H\&K 2011 in following paragraph}

@@

A key choice in devising a multilevel scheme for VIs is how to construct bounds for the coarser levels.  Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} derived from an admissible finest-level iterate.  The next Example illustrates why we do \emph{not}\footnote{An exception occurs in FMG; see Section \ref{sec:implementation}.} apply injection operators directly to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.

\begin{example}  \label{ex:directRbad}  
Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).  Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ strictly less than the maximum of $\underline{\gamma}^1$.  Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.  For appropriate boundary conditions the fine-level constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.  However, various schemes to coarsen the constraints by direct transfers of $\underline{\gamma}^1,\overline{\gamma}^1$ are doomed to failure.  Suppose we propose
\begin{equation}
    \label{eq:fe:badstandard}
    \underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}
\end{equation}
Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.  In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($w^0\in\mathcal{K}^0$; $\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level ($Pw^0 \notin \mathcal{K}^1$).  Attempting to fix this by projecting (truncating) into $\mathcal{K}^1$ would introduce high frequencies not present in $w^0$.  On the other hand, if
\begin{equation}
    \label{eq:fe:badmonotone}
    \underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1 \qquad \text{\emph{[problematic]}}
\end{equation}
then the coarse-level constraint set $\mathcal{K}^0$ is empty because $\underline{\gamma}^0 > \overline{\gamma}^0$ (Figure \ref{fig:directRbad}, right).
\end{example}

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\input{tikz/directRbadfine.tex}
&
\input{tikz/directRbadcoarse.tex}
\end{tabular}
\caption{For the finer-level constraints $\underline{\gamma}^1$ (blue) and $\overline{\gamma}^1$ (red) on the left, direct application of $\iR,\minR,\maxR$ to generate coarser-level constraints is problematic.  Possibility \eqref{eq:fe:badmonotone}, which generates an empty coarser-level set $\mathcal{K}^0$, is at right.  Defect constraints \eqref{eq:fe:defectconstraints} avoid this difficulty.}
\label{fig:directRbad}
\end{figure}

@@

This construction of multilevel box constraints, likewise the unilateral construction in \cite{GraeserKornhuber2009}, permits large, yet admissible, coarse corrections, which promotes multilevel solver efficiency.  When coarsening the problem downward in a V-cycle, one cannot predict the upcoming coarser corrections and so one must construct constraint sets from which any sum will be admissible.  With this in mind, and looking forward to telescoping sums \eqref{eq:fe:telescoping}, we compute the LDC differences \cite{GraeserKornhuber2009}:
\begin{equation}
\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \qquad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}.  \label{eq:fe:philevels}
\end{equation}
(Take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  By definition, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and etc.)  Note that while $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, they are not ordered as in \eqref{eq:fe:chiordering}.

@@

The V-cycle solver in the next Section (Algorithm \ref{alg:fascd}) will compute corrections from $\mathcal{D}^j$ and $\mathcal{U}^j$.  For general box constraints the upward direction is, however, based on \emph{incomplete} CDs as we now explain.  Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, though this becomes a true CD satisfying \eqref{eq:subspacedecomp}--\eqref{eq:decompositionstability} in unilateral cases.  The following Lemma, apparently new, extends the unilateral CD construction underlying Algorithm 4.2 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.

@@

In unilateral cases we can also construct decomposition operators $\Pi_i$ and thereby show \eqref{eq:constraintrestrictionsum}.  For concreteness suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels; the $\underline{\gamma}^J=-\infty$ case is similar.  For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus\,:\,\tilde{\cV}^j \to \tilde{\cV}^i$ be the operator applying the minimum (monotone) injection $j-i$ times: $I_{j\to i}^\ominus v = \minR(\cdots(\minR v)\cdots) $.  (Set $I_{j\to j}^\ominus v = v$ and $I_{j\to -1}^\ominus=0$ by definition.)  Now define nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
\begin{equation}
\Pi_i z^j \coloneqq I_{j\to i}^\ominus(z^j - \underline{\chi}^j) - I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j) + \underline{\phi}^i.  \label{eq:fe:unilateraldecompositionoperator}
\end{equation}
(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property \eqref{eq:fe:monotoneproperty} implies $I_{j\to i}^\ominus(z^j - \underline{\chi}^j) \ge I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j)$, thus $\Pi_i z^j \ge \underline{\phi}^i$, so $\Pi_i z^j \in \mathcal{D}^i$, and furthermore stability \eqref{eq:decompositionstability} holds \cite[Theorem 4.2]{GraeserKornhuber2009}.  On the other hand, the following sum telescopes, leaving only its first term plus the sum in \eqref{eq:fe:telescoping}:
\begin{align*}
\sum_{i=0}^j \Pi_i z^j &= z^j - \underline{\chi}^j + \sum_{i=0}^j \underline{\phi}^i = z^j.
\end{align*}
This shows \eqref{eq:constraintrestrictionsum}, thus \eqref{eq:constraintdecomp}, and that \eqref{eq:fe:downwardsuminclusion} is equality.
\end{proof}

One might attempt to convert \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by constructing different decomposition maps $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.  The above proof shows that the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ indeed holds.  However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.  In fact, $\Pi_i$ defined above does not generally map all non-positive elements of $\mathcal{U}^J$ into $\mathcal{D}^i$ if a finite upper LDC is present.  The Example does not exclude other strategies for showing full CDs in the general box-constrained case.

\begin{example}  \label{ex:notfullcd}
Let $a > 0$ be an arbitrary positive number.  Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.  Note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.  Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse vertices and value $0$ on the other vertices.  Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^1 \in \mathcal{U}^1$.  However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.  From \eqref{eq:fe:unilateraldecompositionoperator}, $\Pi_1 z^1$ simplifies to $\Pi_1 z^1 = z^1 - \underline{\chi}^1$.  But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.  This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, and $\Pi_1 z^1 \notin \mathcal{D}^1$ would then occur.
\end{example}

@@

\begin{lemma}  \label{lem:upwardadmissibility}  For any $j=0,1,\dots,J$ and $0\le k\le j$,
\begin{equation}
\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^{k+1} + \mathcal{U}^k \label{eq:fe:upwardsuminclusion}
\end{equation}
In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, actually $\mathcal{U}^j = \mathcal{U}^k + \sum_{i=k+1}^j \mathcal{D}^i$ is a full CD. \end{lemma}

\begin{proof}  Definition \eqref{eq:fe:philevels} shows inclusion \eqref{eq:fe:upwardsuminclusion}.  In unilateral cases we may follow the proof of Lemma \ref{lem:downwardadmissibility} with \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but modified in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.  Otherwise the proof goes through as before.
\end{proof}

@@

The V-cycle in this Section extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.2]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}; it reduces to FAS for PDEs when inequality constraints are removed.  Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$, via information from each coarser FE subspace $\mathcal{V}^j$, by approximately solving VI problems in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.  The corrections are inductively constructed so that iterates remain admissible at every level.  A key idea is that up-smoothing solves less-restricted VI problems than down-smoothing.  Going downward one does not know what coarse corrections are coming, thus one must choose smaller constraint sets ($\mathcal{D}^j$) such that the correction sum is admissible once we return to a given level.  Going upward, however, no further coarse corrections can violate admissibility, so we relax in the larger sets $\mathcal{U}^j$.

Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdcycles}, left).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Relaxation then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for relaxation on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using inclusion \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.  The new finest-level iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\input{tikz/vcycle.tex}
&
\input{tikz/fcycle.tex}
\end{tabular}
\caption{Left: The FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets. Right: FMG Algorithm \ref{alg:fascd-fmg} generates the initial iterate on finer levels by injection and truncation (doubled edges).}
\label{fig:fascdcycles}
\end{figure}

We must specify the problems solved at each mesh level.  As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.  On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:vi}.  At the start of each V-cycle we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); they do not depend on the coarse-level corrections.

As we descend we define the initial $j$th-level iterates
\begin{equation}
w^j = \iR(w^{j+1} + y^{j+1}),  \label{eq:fe:definew}
\end{equation}
from which we define $\ell^j \in (\cV^j)'$:
\begin{equation}
\ell^j = \begin{cases} \ell^J, & j=J \\
                       f^j(w^j) + R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
\end{equation}
As is typical of FAS \cite[section 5.3]{Trottenbergetal2001}, injection $\iR$ coarsens the (primal) iterates, but canonical restriction $R$ coarsens (dual) linear functionals.  Descending through the levels $j=J,J-1,\dots,1$ we solve the following VI problems for $y^j \in \mathcal{D}^j$:
\begin{equation}
\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j. \label{eq:fe:downvi}
\end{equation}
The initial iterate is $y_{(0)}^j=0$; this is admissible because $\underline{\phi}^j \le 0 \le \overline{\phi}^j$.

@@

However, because the down-smoother has made progress, $y$ should be approximate-able by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the relaxed residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.  These steps yield a problem for the $j$th-level correction $y=y^j \in \mathcal{D}^j$:
\begin{align}
&\ip{f^j\left(\iR(w^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(w^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right)}{v-y^j}, \notag
\end{align}
for all $v\in \mathcal{D}^{j}$.  Though in cluttered form, this is VI problem \eqref{eq:fe:downvi}, once definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} are applied.
